---
title: "Take Home 6"
subtitle: "Winter 2022"
output: pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This homework is due on Tuesday March 8 by midnight PST.

\textbf{Instructions}: Follow each one diligently. \textbf{No exceptions.}
\hrule
\begin{enumerate}
	\item Please type your answers and upload as a PDF in Gradescope. If you prefer to write your answers, we will accept handwritten work provided it is legible. If you decide to go this route, be sure to \href{https://www.camscanner.com)}{scan} and upload a single file.  
	\item Write your full name (as it appears in the grade book) on the first page, preferably as a running header on each page.
	\item Answer the questions in order and number them as I have here. 
	\item \textbf{Write complete answers which includes justifications for steps and/or references to results you use. This is really important.You all need to practice writing mathematical answers - define notation, explain what you are doing, cite any rules, etc. }
	\item Clearly indicate your final answer. Put a frame around it where appropriate.
\end{enumerate}
\hrule 

1. Let us assume that the life of a tire (in miles), say $X$, is normally distributed with mean $\mu$ and standard deviation $\sigma$ \emph{known} to be 5,000 miles. Past experience indicates that $\mu=30,000$. The manufacturer claims that the tires made by a new process have mean $\mu > 30,000$ and it is very possible that $\mu = 35,000$. 

   Let us check their claim by testing $H_0: \mu \leq 30,000$ versus $H_1:\mu > 30,000$. We shall observe a large number $n$  independent values of $X$ say $X_1, X_2,\dots,X_n$ and we shall reject the null hypothesis if $\bar{X} \geq k$.
   
   Determine $n$ and $k$ so that the significance level $\alpha=0.01$ and the type II error rate at $\mu=35,000$ is no more than  1.79\%.
	
2. In this problem, you will write a function called `z.test` which calculates the p-value for testing the mean of a population. You will then use this function repeatedly to calculate the p-value in many randomly samples drawn from various distributions. 

a. Fill in the blanks in the code below for the function `z.test`. The function takes as its arguments 
 - $x:$ an $n \times 1$ vector of data
 - alternative: one of greater, less or two.sided
 - $\mu_0$: the null hypothesized value for the population mean 
 - $\sigma$: the known value of the population standard deviation 
 
We are going to write the function very generally so it calculates the p-value for one sided and two sided alternatives. 
When you are done, remove the chunk option  `eval = FALSE` to reveal and evaluate the completed code in the knitted output.

```{r eval = FALSE}

z.test <- function( x, alternative = c("greater", "less", "two.sided"), mu0 = 0, sd = 1){
  
  alternative = match.arg(alternative) #fancy matching
  
  n <- length(x)    # number of observations
  xbar <- ___       #mean of x
  z <- ___          # z statistic for testing H0: mu = mu0
  
  #calculate p-value
  pvalue = ifelse( alternative == "two.sided", ___, 
                   ifelse(alternative == "greater", ___, ___))
  
  return(pvalue)
}

z.test(x = rnorm(n = 25, mean = 0, sd = 1), 
       alternative = "two.sided",
       mu0 = 0, 
       sigma = 1)

```


b. Create a dataframe called `null_sim` which calculates the p-value for testing $H_0: \mu = 5$ versus $H_1: \mu \neq 5$ in 1,000 samples each of size $n = 25$ from a $Norm(5, 1)$ distribution.   (\emph{Hint:} use the `replicate` function. See the R exercise following example 3.2 in the chapter titled "Hypothesis Tests: p-values and error probabilities" for an idea of how to get started.)

c. Make a QQplot to compare the quantiles of the p-values with those of a $Unif(0, 1)$ distribution. Does the uniform seem like a good fit?

d. Repeat parts b. and c. but this time, simulate 1,000 samples of size $n=10$ from a $Norm( 6, 1)$ distribution. Does the uniform still seem like a good fit for the p-values?


3. A manufacturer of toothpicks wonders what the mean width of a toothpick is under a new
manufacturing method. How many toothpicks must the manufacturer measure to be 90\% confident that
the sample mean is no farther from the true mean than 0.10 mm? Assume the conditions of the Central Limit Theorem are satisfied, and note that
toothpicks produced under the old method had a standard deviation of 0.4 mm.

4. For any random variable $X$ with mean $\mu$ and standard deviation $\sigma$, Chebyshev's inequality states that
$$P( | X - \mu| \geq k \sigma ) \leq \frac{1}{k^2},$$
where $k$ is some real number. Suppose $\widehat{\theta}$ is an unbiased estimator of $\theta$ with standard error $\sigma(\widehat{\theta})$. (For example if $\widehat{\theta} = \bar{X}$, then $\sigma(\bar{X}) = \frac{\sigma}{\sqrt{n}}$.)

   Use Chebyshev's inequality to construct a confidence interval which has at least 95% coverage probability.