---
title: "Problem Section 3 KEY"
subtitle:  "Combining Random Variables"
graphics: yes
output: pdf_document
header-includes:
    - \usepackage{amsmath, amssymb}
    - \usepackage{framed}\definecolor{shadecolor}{rgb}{0.949,0.949,0.949}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

* * * 

\begin{shaded}

\textbf{Learning Outcomes}

The problems are designed to build conceptual understanding and problem-solving skills. The emphasis is on learning to find, evaluate and build confidence. The specific tasks include: 

 
   - Finding the distribution of the sum of two independent random variables
   
  - Finding the mean and variance of the sum of independent normal random variables.
  
   - Performing probability calculations with a normal distribution
   
   - Be able to work out probabilities for the maximum of two continuous random variables 

   - Back up and support work with relevant explanations


\end{shaded}


* * *

### Exercises 

1. We can write the joint PMF of $\langle X, Y \rangle$ in a tabular form. Each $(x,y)$ pair has probability $\frac{1}{9}$ since the random variables are independent. 

\begin{table}[h]
\centering
\begin{tabular}{c|ccc}
& \multicolumn{3}{c}{$x$} \\ \hline
$y$ & 0 & 1 & 2  \\ \hline
0 & 1/9 & 1/9 & 1/9  \\
1 & 1/9 & 1/9 & 1/9  \\
2 & 1/9 & 1/9   & 1/9 
\end{tabular}
\end{table}

We have that $S$ can take the following 5 values, $s=0,1,2,3,4$.
Thus we have the following PMF for $S$:

\begin{table}[h]
\centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline
 $s$ & 0 & 1 & 2 & 3 & 4 \\ \hline
 outcomes & (0,0) & (0,1); (1,0) & (0,2), (1,1), (2,0) & (1,2), (2,1) & (2,2)\\
$f(s)$ & $\frac{1}{9}$ & $\frac{2}{9}$ & $\frac{3}{9}$ & $\frac{2}{9}$ & $\frac{1}{9}$ \\ \hline
\end{tabular}
\end{table}


2.

a. Let $X_1, \dots, X_5$ denote the number of cars crossing the bridge on the 5 weekdays. More specifically we are given that:
$$
X_i \overset{ind}{\sim} N(\mu = 67000,\sigma = 1000) \ , i=1,2,3,4,5
$$

Similarly let $Y_1,Y_2$ be the number of cars crossing on the weekends, so we are given that:
$$
Y_j \overset{ind}{\sim} N(\mu = 72000,\sigma = 1000) \ , j=1,2
$$

We wish to find the distribution of $T = X_1 + X_2 + \dots + X_5 + Y_1 + Y_2 = \sum_{i=1}^5X_i + \sum_{j=1}^2 Y_j$.

Since these are independent random variables, by the properties of the normal distribution, we know T will be normally distributed. We will have that:

\begin{align*}
E(T) &= E(\sum_{i=1}^5X_i + \sum_{j=1}^2 Y_j) \\
&= \sum_{i=1}^5 E[X_i] + \sum_{j=1}^2 E[Y_j] \ \text{by linearity of expectation} \\
&= 5E[X_1] + 2E[Y_1] \ \text{since each X and Y follows the same distribution} \\
&= 5\times 67000 + 2 \times 72000 \\
&= `r 5*67000 + 2*72000` 
\end{align*}

Similarly we will have that:

\begin{align*}
Var(T) &= Var(\sum_{i=1}^5X_i + \sum_{j=1}^2 Y_j) \\
&= \sum_{i=1}^5 Var[X_i] + \sum_{j=1}^2 Var[Y_j] \ \text{by independence of the RV's} \\
&= 5Var[X_1] + 2Var[Y_1] \ \text{since each X and Y follows the same distribution} \\
&= 5\times 1000^2 + 2 \times 1000^2 \\
&= 7000000
\end{align*}

Thus we have that $SD(Z) = `r sqrt(7000000)`$

Thus we have that the total number of cars crossing the bridge in a 7 day week, W, has distribution:

$$
Z \sim N(\mu = `r 5*67000 + 2*72000` ,\sigma = \sqrt{7000000} = `r sqrt(7000000)`)
$$


b. We wish to find $P(T>482,000)$. We can simply use pnorm to find this in R:

```{r}
p_t_482 <- 1 - pnorm(482000,mean = 5*67000 + 2*72000, sd = sqrt(7000000))
```

Thus we see that $P(T>482,000) = `r p_t_482`$


3. 

$$
P(max(X_1,X_2)>q) = 1 - P(max(X_1,X_2)\leq q) =  1 - P(X_1 \leq q \ \cap X_2\leq q)
$$

Since $X_1, X_2$ are independent RV's, we have that $P(X_1 \leq q \ \cap X_2\leq q) = P(X_1 \leq q) \times P(X_2 \leq q)$.

Since q is the 60th percentile, we know that $P(X_1 \leq q) = P(X_2 \leq q) = .6$. Thus we have that:

$$
P(X_1 \leq q \ \cap X_2\leq q) = P(X_1 \leq q) \times P(X_2 \leq q) = .6 \times .6 = .36
$$

Thus putting it all together we have:

$$
P(max(X_1,X_2)>q) = 1 - P(max(X_1,X_2)\leq q) =  1 - P(X_1 \leq q \ \cap X_2\leq q) = 1-.36 = .64
$$
So the probability that the larger of the 2 RV's is greater than q is .64   
   

4. 

a. Find the PDF of $X_{max} = \max\{X_1,X_2,X_3\}$

We know by Theorem 16.5 that $f_{max}(x) = n[F(x)]^{n-1}f(x)$. For an exponential RV, we have that $f(x) = \lambda exp(-\lambda x)$ and that $F(x) = 1-e^{-\lambda x}$. In this case we have that n=3 and $\lambda=1$. 

Thus we have that:

$$
f_{max}(x) = n[F(x)]^{n-1}f(x) = 3[1-e^{-x}]^{3-1} e^{-x} \ \ \ x \geq 0
$$

Doing some simplification we have that:

\begin{align*}
f_{max}(x) &= 3[1-e^{-x}]^{2} e^{-x} \\
&= 3 e^{-x}[1-2e^{- x} + e^{-2 x}] \\
&= 3 e^{-x} - 6 e^{-2 x} + 3 e^{-3 x} \ \ x \geq 0
\end{align*}


b. 

```{r echo = TRUE}

pdf_max <- function(x){
  #return the PDF of Xmax at a given value x
  3*exp(-x)-6*exp(-2*x)+3*exp(-3*x)
}


ggplot(data = NULL) +
  #first layer is for X ~ Exp(1)
  geom_function(fun = dexp, 
                args = list(rate = 1),
                mapping = aes(color = "Exp(1)" ),
                xlim=c(0,10) ) +
  #second layer is for PDF of X_{max}
  geom_function(fun = pdf_max,
                mapping = aes(color = "Xmax"),
                xlim=c(0,10)  ) +
  labs(x = "x",
       y = "density",
       title = "Comparing distributions of X and Xmax",
       color = "Dist")
  

```
