---
title: "Problem Section 1 KEY"
subtitle: "Discrete Distributions"
graphics: yes
output: pdf_document
header-includes:
    - \usepackage{amsmath, amssymb}
    - \usepackage{framed}\definecolor{shadecolor}{rgb}{0.949,0.949,0.949}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\begin{shaded}

\textbf{Learning Outcomes}

The problems are designed to build conceptual understanding and problem-solving skills. The emphasis is on learning to find, evaluate and build confidence. The specific tasks include: 

   - Construct a table of joint probabilities and calculate marginal and conditional distributions.
     
   - Calculate probabilities for the trinomial experiment
  
   - Back up and support work with relevant explanations


\end{shaded}


* * *


### Exercises 

1. A fair coin is tossed four times. $X$ is the number of heads that come up on the first three tosses and $Y$ is the number of heads that come up on tosses 2, 3, 4. 

a. Fill in the cells of the table showing the joint PMF of $\langle X, Y \rangle$. 
   
   \begin{table}[h]
   \centering
   \caption{Joint and marginal PMF of $X$ and $Y$}
   \begin{tabular}{|c|c|c|c|c|c|}
   \hline
    & x=0 & x=1 & x=2 & x=3 & $f_Y(y)$\\
   \hline
   y=0 & `r .5^4` & `r .5^3*.5` & 0 & 0 & `r dbinom(0,3,.5)`\\
   y=1 & `r .5^3*.5` & `r .5^4 + .5^2*.5^2*2` & `r .5*.5^2*2*.5` & 0 & `r dbinom(1,3,.5)` \\
   y=2 & 0 & `r .5*.5^2*2*.5` & `r .5^4 + .5^2*.5^2*2` & `r .5^4` & `r dbinom(2,3,.5)`\\
   y=3 & 0 & 0 & `r .5^3*.5` & `r .5^4` & `r dbinom(3,3,.5)`\\
   \hline
   $f_{X}(x)$ & `r dbinom(0,3,.5)` &  `r dbinom(1,3,.5)` & `r dbinom(2,3,.5)` &  `r dbinom(3,3,.5)` & \\ \hline
   \end{tabular}
   \end{table}
   
    \textbf{Explanation:} By the multiplication rule for counting, there are $$2 \times 2 \times 2 \times 2 = 2^{4} = 16$$ possible outcomes in this sample space. Each outcome is equally likely and therefore the probabilities of the various $(x,y)$ combinations can be deduced by simply counting the number of outcomes that would result in this value.
   
    For example, the outcomes that lead to $x=1, y=2$ are HTTH, THTH, TTHH, which gives a probability of 3/16.

   
   
b. Find the marginal distribution of $X$ and also of $Y$.
   
    We can find the marginal distributions by simply adding the probabilities along the rows or the columns. However, we can also recognize that since $X$ is the number of heads in 3 tosses, it is a binomial random variable. Hence it's probabilities are given by the $Binom(3, \frac{1}{2})$ distribution. 
    
    The marginal probabilities are shown in the table in part a along the margins. 
    
   
c. Find the conditional probabilities $P(X =2 | Y = 2)$ and $P(Y = 3|X = 0)$. 

    Using the formula for conditional probabilities
    $$P(X = x|Y=y) = \frac{ P(X = x, Y = y)}{P(Y = y)}$$
    we have 
    $$P(X = 2|Y = 2) = \frac{0.1875}{0.375} = `r 0.1875/0.375`$$
    and 
    $$P(Y = 3|X = 3) = \frac{0}{0.125} = 0$$

2. Suppose that five independent observations are
drawn from the continuous PDF
    $$f(t) = 2t \ \ 0 \leq t < 1$$
Let $X$ denote the number of $t$'s that fall in the interval $0 \leq t < \frac{1}{3}$ and let $Y$ denote the number of $t$'s in the interval $\frac{1}{3} \leq t < \frac{2}{3}$. 

a. Find the probability that $X=1$ and $Y = 2$.

   Here $\langle X, Y \rangle$ follow a trinomial distribution since 
   - there are 5 (fixed number) trials
   - each trial results in one of three outcomes: whether the realization falls in $[0, \frac{1}{3})$, or $[\frac{1}{3}, \frac{2}{3})$ or $[\frac{2}{3}, 1)$.
   - the trials are independent 
   
     Under these conditions, the joint PMF of $\langle X, Y \rangle$ is
  \begin{align*}
   f(x,y) &= P(X = x, Y = y), \\
          &= \frac{5!}{x!y! (5-x-y)!} \pi_1^{x} \pi_2^{y} (1-\pi_1-\pi_2)^{5-x-y}
  \end{align*}
  
    where $x+y \leq 5$ and the probabilities $\pi_1$ and $\pi_2$ are
  \begin{align*}
  \pi_1 &= \int\limits_{0}^{1/3} f(t) dt = \frac{1}{9},\\
  \pi_2 &= \int\limits_{1/3}^{2/3} f(t) dt = \frac{3}{9}.
  \end{align*}
  
    We will calculate the probability that $X=2, Y=1$ by plugging into the joint PMF:
  \begin{align*}
  f(2,1) &= P(X=2, Y=1), \\
         & \frac{5!}{2! 1! 2!} \left(\frac{1}{9} \right)^{2} \times \left(\frac{3}{9} \right) \:\times \left(\frac{5}{9} \right)^{2}, \\
         &= `r (factorial(5)/4)*(1/9)^2*(3/9)*(5/9)^2`
  \end{align*}
  
    We can also verify this in R.
    ```{r}
    dmultinom(x=c( 2, 1, 2), prob = c( 1/9, 3/9, 5/9))
    ```
  
b. Find the probability that $X=1$. How will you calculate this in R?

    Since this is a statement just about $X$, we will need to use the marginal distribution of $X$ in order to calculate the probability. 
    
    We know that the marginal distribution of $X$ is a binomial. In other words: $X \sim Binom( 5, \frac{1}{9})$. We want $P(X=1)$.
    
    ```{r}
    dbinom(x = 1, size = 5, prob =1/9 )
    ```

c. How many observations should you expect will lie in the interval $0 \leq t < \frac{1}{3}$. That is, find $E\left[X \right]$. What is the standard deviation of $X$?

    Marginally $X \sim Binom( 5, \frac{1}{9})$. Therefore, 
    \begin{align*}
    E\left[X \right] &= \frac{5}{9},\\
    SD\left[X \right] &= \sqrt{ 5 \times \frac{1}{9} \times \frac{8}{9} } = \frac{2 \times \sqrt{10}}{9}.
    \end{align*}

    We should expect `r round(5/9,4)` values to fall in the  interval $0 \leq t < \frac{1}{3}$,  give or take `r round(sqrt(5*8/81),4)`. 



3. (For HW 1 problem 3) The binomial theorem states that for real numbers $a$ and $b$ and an integer $n$:

    $$(a+b)^{n} = \binom{n}{0} a^{0} b^{n} + \binom{n}{1} a^{1} b^{n-1} + \binom{n}{2} a^{2} b^{n-2} + \dots + \binom{n}{n} a^{n} b^{0}$$
     
     Show, using the binomial theorem, that
$$\sum\limits_{x=0}^{n} \binom{n}{x} = 2^n.$$ 

     The binomial theorem can be written succinctly as
     $$(a+b)^n = \sum\limits_{x=0}^{n} \binom{n}{x} a^{x} b^{n-x}$$
     
     Choosing $a=1$ and $b=1$ we then get the result
     $$(1+1)^n = 2^n = \sum\limits_{x=0}^{n} \binom{n}{x}.$$